---
title: "PA4 🙂"
subtitle: "Eva Corregidor"
author: "Data Science for Linguists"
institute: "Spring 2025"
date: "2025/03/31 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
library(here)
library(ggthemes)
```
# First Steps
1 - Create a new project in RStudio called pa4. Inside your new project, create three folders: data_raw, data_tidy, and slides.

2 - Download the dataset available and store it in data_raw.

3 - Generate an HTML presentation using xaringan. Save the RMarkdown file as index.Rmd inside of the folder slides.
---
# Step 4
Load the dataset from inside your index.Rmd file.
```{r}
pa4_data <- read.csv(here("data_raw", "fricatives", "fricatives.csv"))
```
---
# Step 5
Tidy the data and save the tidy version as a .csv file in data_tidy.
```{r}
pa4_tidy <- pa4_data |>
  pivot_longer(cols = -obs, names_to = "measure", values_to = "value") |> 
  separate(measure,into = c("phoneme", "measurement"), sep = "_")|>
  pivot_wider(names_from = measurement, values_from = value)

write.csv(pa4_tidy, "tidy_data.csv", row.names = FALSE)
```

---
# Step 6
Provide a table of descriptive statistics.
```{r}
#| echo: false

pa4_tidy |> 
  group_by(phoneme) |> 
  summarize(avg = mean(cog), sd = sd(cog)) |>
  kable()
```
---
# Step 7: Boxplot
Make a boxplot of center of gravity as a function of phoneme. Use a statistical transformation (i.e., not a boxplot, but rather stat_summary()).
```{r}
#| echo: false
#| fig-asp: 0.56
#| out-width: "100%"
#| dpi: 300
pa4_tidy |>
  ggplot(aes(x = phoneme, y = cog, color = phoneme)) +
  geom_jitter(height = 0, width = 0.2, alpha = 0.4) +
  stat_summary(
  fun.data = mean_sdl,
  geom = 'pointrange'
) +
  scale_color_viridis_d(guide = "none", begin = 0.2, end = 0.8) +
  labs(
    title = "Center of gravity as a function of phoneme",
    x = "Phoneme: [s] or [ʃ]", y = "Center of gravity")
```
---
# Step 7: Skewness
In another slide, plot skewness as a function of phoneme. Use a statistical transformation (i.e., not a boxplot, but rather stat_summary()).
```{r}
#| echo: false
#| fig-asp: 0.56
#| out-width: "100%"
#| dpi: 300
pa4_tidy |>
  ggplot(aes(x = phoneme, y = skewness, color = phoneme)) +
  geom_jitter(height = 0, width = 0.2, alpha = 0.4) +
  stat_summary(
  fun.data = mean_sdl,
  geom = 'pointrange'
) +
  scale_color_viridis_d(guide = "none", begin = 0.2, end = 0.8) +
  scale_fill_viridis_d(guide = "none", begin = 0.2, end = 0.8) +
    labs(
    title = "Skewness as a function of phoneme",
    x = "Phoneme: [s] or [ʃ]", y = "Skewness")
```
---
# Step 8
Fit a model that examines center of gravity as a function of skewness for the [s] segments (hint: you will have to transform the data). Make a table of the model summary.

First, I transform the data.
```{r}
pa4_transformed <- pa4_data |>
  pivot_longer(cols = -obs, names_to = "measure", values_to = "value") |>
  separate(measure,into = c("phoneme", "measurement"), sep = "_") |>
  filter(phoneme == "s")|>
  pivot_wider(names_from = measurement, values_from = value)
```
---
# Step 8: Cont.
Second, I fit the model and make a table of it's summary.
```{r}
#| echo: false
pa4_model <- lm(cog ~ skewness, data = pa4_transformed)

summary(pa4_model)$coef |> 
  kable()
```
---
# Step 9
Make a scatter plot that illustrates the relationship in (8).
```{r}
#| echo: false
#| fig-asp: 0.56
#| out-width: "100%"
#| dpi: 300
pa4_model |>
  ggplot() +
  geom_point(aes(x = cog, y = skewness), color = "blue") +
    labs(
    title = "Center of gravity as a function of skewness for the [s] segments",
    x = "Center of gravity", y = "Skewness")
```
---
# Step 10
Check model diagnostics (make plots).
```{r}
#| echo: false
#| fig-asp: 0.56
#| out-width: "100%"
#| dpi: 300
ds4ling::diagnosis(pa4_model)
```
---
# Step 11
Write up the results (as if it were for a journal article).

For this PA4, I fitted a linear model to explain the relationship between the center of gravity (cog) and skewness of the [s] segments in our data. The model revealed a main effect of skewness (F(1, 48) = 144.5, p<0.001) and provided the best fit of the data (R-squared = 0.75). The visual inspection of Q-Q plots and plots of residuals (see slides #10 and #11) revealed no discernable deviations from normality and homoskedasticity. Overall, the center of gravity increased as a function of skewness.
---
# Step 12
In a new slide, load the assumptions.csv dataset. 
```{r}
pa4_assumptions <- read.csv(here("data_raw", "fricatives", "assumptions.csv"))
```
---
# Step 12: Cont.
Make a scatterplot.
```{r echo=FALSE}
pa4_assumptions |>
  ggplot() +
  geom_point(aes(x = x, y = y), color = "red")
```
---
# Step 12: Cont.
Explain in a few sentences why it would not be appropriate to fit a model to this data.

Linearity is not met. Deviations with such a strength visually indicate that a linear model is not a great fit for this data set. It would be really hard to interpret the relationship of x and y.
---
# Last Steps
13 - Host your project in a GitHub repo called pa4.

14 - Turn the slides into a website using GitHub pages.

15 - Inside your dropbox in programming_assignments, create a folder called pa4 that includes a README.md file with a link to your slides.

16 - Push changes to your forked version of programming_assignments and submit a pull request to the master programming_assignments repo in ds4ling.
